---
name: 'Mojisoluwa (Jiso) Awe'
title: '2019 WAR Visualizations'
output: html_notebook
---

```{r}  
library(tidyverse)     ##loading packages
library(ggplot2)
library(dcData)
library(rvest)
library(dplyr)
library(lubridate)
```

#Wins Above Replacemnt?

In this notebook we will be examining the advanced statistic of Wins Above Replacement (WAR) from the 2019 MLB season. 

A simple explanation of the WAR statistic is a number defining the amount of many wins (or loss) a player contributes to a teams record. According to Baseball Reference WAR can be defined as 'a single number that represents the number of wins the player added to the team, above what a replacement player would add'.

We will examining players WAR from the 2019 season, this was the most recent 162 game season we have data from. This data has been scraped from ESPN.com.

```{r}
page <- 'https://www.espn.com/mlb/war/leaders/_/type/seasonal/year/2019'
tableA <- page %>%
  read_html() %>%
  html_nodes(css = 'table') %>%
  html_table(fill = TRUE)

tableA
```

```{r} 
##creating the right data frame
WarT <-
  tableA[[1]]

WarT
```

```{r}
##removing the variables from the variables from the dataframe that aren't going to be used
WarT <-
  subset(WarT, select = -c(X6:X10))
WarT
```



```{r}
##renaming variables and removing header cases
wt <-
  WarT%>%
  filter(row_number() > 2) %>%
  rename( Rank =  `X1` ,
         Player =  `X2` ,
         WAR =  `X3`,
         OFF = `X4`,
         DEF = `X5`)

wt
```











